Trends of SDR in Thessaloniki

### 2023-11-24 Article_A.tex
Commit 2023-11-24 14:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  preprint, 3p, authoryear]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Trends of SDR in Thessaloniki},
  pdfkeywords={GHI; SDR; Solar Brigthening/Dimming.},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Trends of SDR in Thessaloniki}
\author{true \and true}
\date{2023-11-03}

@@

For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
To calculate the reference clear sky
\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\) derived by
the radiation model of Haurwitz (1945), adjusted for our site with a
factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative optimization process, as described
by Long and Ackerman (2000) and Reno and Hansen (2016).
The target of the optimization was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and
was accomplished with the algorithmic function ``optimise'', which is an implementation based on the work of Brent (1973), from the library ``stats'' of the R programming language (R Core Team 2023).
The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by Reno, Hansen, and Stein (2012b) and evaluated by Reno and Hansen (2016).
We found, that Haurwitz's model, adjusted with the factor \(a = 0.965\) yields one of the lowest root mean squared errors (RMSE),
while the procedure, manages to characterize the majority of the data.
Thus, our clear sky reference is derived by the Eq.~\ref{eq:ahau}.
\begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation}
where: \(n\) is the total number of daylight records, \(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky by CSid, \(a\) is a hypothetical adjustment factor, and \(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky radiation models.
\begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation}
where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in \(\text{W}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle (SZA).

@@

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\tightlist
\item
  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq. \ref{eq:MeanVIP}).
  \begin{equation}
  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20\,Wm^{-2}
  < \overline{\text{SDR}}_i <
  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30\,Wm^{-2}
  \label{eq:MeanVIP}
  \end{equation}
\item
  Maximum measured value \(M_{i}\) (Eq.~\ref{eq:MaxVIP}).
  \begin{equation}
  1 \times M_{\text{CSref},i} - 75\,Wm^{-2}
  < M_{i} <
  1 \times M_{\text{CSref},i} + 75\,Wm^{-2}
  \label{eq:MaxVIP}
  \end{equation}
\item
  Length \(L_i\) of the sequential line segments, connecting the points of the \(11\) SDR values (Eq. \ref{eq:VILeq}).
  \begin{equation}
  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
  \label{eq:VILeq}
  \end{equation}
  \begin{equation}
  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
  \label{eq:VILcr}
  \end{equation}
  where: \(t_i\) is the time stamp of each SDR measurement.
\item
  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the \(11\) sequential points, normalized by the mean \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}).
  \begin{gather}
    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
  \end{gather}
  For this criterion, \(\sigma_i\) should be below a certain threshold (Eq.~\ref{eq:VCTcr}):
  \begin{equation}
    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
  \end{equation}
\item
  Maximum difference \(X_i\) between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
  \begin{gather}
    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
  \end{gather}
  For this criterion, \(X_i\) should be below a certain threshold (Eq.~\ref{eq:VSMcr}):
  \begin{equation}
    X_i < 7.5 \label{eq:VSMcr}
  \end{equation}
\end{enumerate}

Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and buildings), we excluded data with azimuth angle between
\(35^\circ\) and \(120^\circ\) with SZA greater than \(80^\circ\).
On the latter instances, Sun is systematically not visible from the instrument's location.
To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
The TSI data we used are part of the ``NOAA Climate Data Record of Total Solar Irradiance'' dataset (Coddington et al. 2005).
The initial daily values of this dataset were interpolated to match the time step of our measurements.
In the final dataset
\(6589967\)
one-minute measurements, of which,
\(84.2\%\)
were identified as under clear-sky conditions and
\(15.8\%\)
as under cloud-sky conditions.

@@

We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table~\ref{tab:trendtable}).
In Figure~\ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}).
We observe a positive trend for all-sky conditions
(\(0.38\,\%/y\)), a very close but lower trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend for cloud-skies (\(-0.28\,\%/y\)).
In the studied period, there is no significant break or change in the variability of the time series.
Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records (Wild et al. 2021; Yuan, Leirvik, and Wild 2021; Ohmura 2009). It is interesting to note, that for the observation period, the trend of the TSI is
\(-0.0002\,\%/y\), and thus we can not attribute any major effect on SDR trend to Solar variability.

@@

Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions (Regier, Briceño, and Boyer 2019).
In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
For cloud-sky (Figure~\ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
In our case, there is more complex evolution of the phenomenon.
Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table~\ref{tab:trendtable}).

@@

The solar zenith angle is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols (Wang et al. 2021).
In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of \(1^\circ\) SZA, and calculated the overall trend for each bin before noon and after noon (Figure~\ref{fig:szatrends}).
Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year (Wang et al. 2021).
For cloud-sky conditions (Figure~\ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
For SZAs \(16^\circ\) - \(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a weak statistical significance.
Between \(50^\circ\) and \(75^\circ\) SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  preprint, 3p, authoryear]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{subcaption}
\captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Trends of SDR in Thessaloniki},
  pdfkeywords={GHI; SDR; Solar Brigthening/Dimming.},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Trends of SDR in Thessaloniki}
\author{true \and true}
\date{2023-11-24}

@@

For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
To calculate the reference clear sky \(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of Haurwitz (1945), adjusted for our site with a factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative optimization process, as described by Long and Ackerman (2000) and Reno and Hansen (2016).
The target of the optimization was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and was accomplished with the algorithmic function ``optimise'', which is an implementation based on the work of Brent (1973), from the library ``stats'' of the R programming language (R Core Team 2023).
The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
During the optimization, eight simple clear sky radiation models were tested (namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by Reno, Hansen, and Stein (2012b) and evaluated by Reno and Hansen (2016).
We found, that Haurwitz's model, adjusted with the factor \(a = 0.965\) yields one of the lowest root mean squared errors (RMSE), while the procedure, manages to characterize the majority of the data.
Thus, our clear sky reference is derived by the Eq.~\ref{eq:ahau}.
\begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation}
where: \(n\) is the total number of daylight records, \(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky by CSid, \(a\) is a hypothetical adjustment factor, and \(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky radiation models.
\begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation}
where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in \(\text{W}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle (SZA).

@@

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\tightlist
\item
  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq. \ref{eq:MeanVIP}).
  \begin{equation}
  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20\,Wm^{-2}
  < \overline{\text{SDR}}_i <
  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30\,Wm^{-2}
  \label{eq:MeanVIP}
  \end{equation}
\item
  Maximum measured value \(M_{i}\) (Eq.~\ref{eq:MaxVIP}).
  \begin{equation}
  1 \times M_{\text{CSref},i} - 75\,Wm^{-2}
  < M_{i} <
  1 \times M_{\text{CSref},i} + 75\,Wm^{-2}
  \label{eq:MaxVIP}
  \end{equation}
\item
  Length \(L_i\) of the sequential line segments, connecting the points of the \(11\) SDR values (Eq. \ref{eq:VILeq}).
  \begin{equation}
  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
  \label{eq:VILeq}
  \end{equation}
  \begin{equation}
  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
  \label{eq:VILcr}
  \end{equation}
  where: \(t_i\) is the time stamp of each SDR measurement.
\item
  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the \(11\) sequential points, normalized by the mean \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}).
  \begin{gather}
    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
  \end{gather}
  For this criterion, \(\sigma_i\) should be below a certain threshold (Eq.~\ref{eq:VCTcr}):
  \begin{equation}
    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
  \end{equation}
\item
  Maximum difference \(X_i\) between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
  \begin{gather}
    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
  \end{gather}
  For this criterion, \(X_i\) should be below a certain threshold (Eq.~\ref{eq:VSMcr}):
  \begin{equation}
    X_i < 7.5\,Wm^{-2} \label{eq:VSMcr}
  \end{equation}
\end{enumerate}

Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and buildings), we excluded data with azimuth angle between
\(35^\circ\) and \(120^\circ\) with SZA greater than \(80^\circ\).
On the latter instances, Sun is systematically not visible from the instrument's location.
To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
The TSI data we used are part of the ``NOAA Climate Data Record of Total Solar Irradiance'' dataset (Coddington et al. 2005).
The initial daily values of this dataset were interpolated to match the time step of our measurements.
In the final dataset
\(84.2\%\)
of the data were identified as under clear-sky conditions and
\(15.8\%\)
as under cloud-sky conditions.

@@

We calculated the linear SDR trends, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table~\ref{tab:trendtable}).
In Figure~\ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}).
We observe a positive trend for all-sky conditions
(\(0.38\,\%/y\)), a very close but smaller trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend for cloudy-skies (\(-0.28\,\%/y\)).
In the studied period, there is no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR slope, around 1980 (Wild et al. 2021; Yuan, Leirvik, and Wild 2021; Ohmura 2009), a few years before the start of our records.
It is interesting to note, that for the observations period, the trend of the TSI is
\(-0.0002\,\%/y\),
and thus we can not attribute any major effect on SDR trend to Solar variability.

@@

Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) (Regier, Briceño, and Boyer 2019) of the anomalies can reveal different structures in the records of all three sky conditions.
In the cases of all-sky and clear-sky conditions (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
A downward part from the start until about 2005, a relatively steady part until about 2016 and, finally, a steep upward part until the present.
For cloud-sky (Figure~\ref{fig:cusummonth-3}), we have a different pattern; it begins with a relatively steady part until 1997, followed by an upward part until 2005, and a long decline until 2020, with a small positive slope until the present.
For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
In our case, there is a more complex evolution of the anomalies.
Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
Although, there seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main driver for this relation due to the great difference in the number of observational data between the two datasets (Table~\ref{tab:trendtable}).

@@

The solar zenith angle is a major factor of SDR reaching the ground, due to the enhancement of the radiation path in the atmosphere, especially in urban environments with human activities emitting aerosols (Wang et al. 2021).
In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of \(1^\circ\) SZA, and calculated the overall trend for each bin before noon and after noon (Figure~\ref{fig:szatrends}).
Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year (Wang et al. 2021).
For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
For SZAs \(16^\circ\) - \(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a weak statistical significance.
Between \(50^\circ\) and \(75^\circ\) SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
